### 2021-06-22 src/paper/appendix_data.tex
Update paper with new results.

\subsection{The Synthetic Population\comment[id=K]{To be written}}

@@

% other contacts
For the last group of contacts which cover things like leisure activities, grocery
shopping etc. we have no reliable data by how much policies reduce them. In addition,
they are likely to be affected by social and psychological factors such as pandemic
fatigue and vacations. Because of this we estimate them like the infection probabilities
to fit the time series data. We use very few change points and tie them to particular
events such as policy announcements or particular holidays. Because of the scarce data
situation we cannot distinguish between a hygiene factor (such as mask wearing) during
meetings and physical distancing (such as virtual meetings with
friends).\comment[id=K]{@Janos: Maybe make more concrete when the estimation is finished
which phases we have and why the switching points are where they are.}

>>>>

\subsection{The Synthetic Population \comment[id=K]{To be written}}

@@

For the last group of contacts, which cover things like leisure activities, grocery
shopping, etc., we have no reliable data by how much policies reduce them. In addition,
they are likely to be affected by social and psychological factors such as pandemic
fatigue and vacations. Because of this we estimate them like the infection probabilities
to fit the time series data. We use very few change points and tie them to particular
events such as policy announcements or particular holidays. Because of the scarce data
situation we cannot distinguish between a hygiene factor (such as mask wearing) during
meetings and physical distancing (such as virtual meetings with
friends).\comment[id=K]{@Janos: Maybe make more concrete when the estimation is finished
which phases we have and why the switching points are where they are.}

### 2021-06-22 src/paper/report.tex
Update paper with new results.

This paper develops a quantitative model incorporating these factors simultaneously. The
framework allows to combine a wide variety of data and mechanisms in a timely fashion,
making it useful to predict the effects of various interventions. We apply the model to
Germany, where new infections fell by almost 80\% during the month of May 2021. Our
analysis shows that frequent and large-scale rapid testing caused the bulk of this
decrease, which is in line with prior predictions \citep{Mina2021}. We conclude that it
should have a large role for at least as long as vaccinations have not been offered to
an entire population.

At the core of our agent-based model are physical contacts between heterogeneous agents
(Figure~\ref{fig:model_contacts_infections}).\footnote{A detailed comparison with other
approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The model
most closely related to ours is described in \citet{Hinch2020}.} Each contact between an
infectious individual and somebody susceptible to the disease bears the risk of
transmitting the virus. Contacts occur in up to four networks: Within the household, at
work, at school, or in other settings (leisure activities, grocery shopping, medical
appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
applications can take the population and household structure from census data and the
network-specific frequencies of contacts from diary data measuring contacts before the
pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network,
\replaced[id=K]{meeting probabilities depend on}{contacts are differentiated by}
age and
geographical location (see Section~\ref{sub:number_of_contacts} of the Supplementary
Material).

@@

    \floatfoot{\noindent \textit{Note:} A        % formal
        description of the model can be found in Supplementary Material~\ref{sec:model}.
        Figure~\ref{fig:model_contacts_infections} shows the influence on an agent's
        contacts to other agents. Demographic characteristics set the baseline number of
        contacts in different networks. She may reduce the number of contacts due to
        NPIs, showing symptoms, or testing positively for SARS-CoV-2. Infections may
        occur when a susceptible agent meets an infectious agent; the probability depends
        on the type of contact, on seasonality, and on NPIs. If infected, the infection
        progresses as depicted in Figure~\ref{fig:model_disease_progression}. If rapid
        tests are available, agents' demand is modeled as in
        Figure~\ref{fig:model_rapid_tests}. \textcolor{blue}{For each reason only a share
        of eligible individuals receive a test. The share of eligible individuals that
        get tested depends on the reason and changes over time (see
        Figure~\ref{fig:rapid_test_demand}).} \textcolor{red}{All pupils are attending
        school are tested, other reasons trigger a test only for a fraction of
        individuals.} Figure~\ref{fig:model_official_cases} shows the model of
        translating all infections in the simulated data to age-specific recorded
        infections. The model uses data on the aggregate share of recorded cases, on the
        share of positive PCR tests triggered by symptoms, \textcolor{red}{and on the
        false positive rate of rapid tests}. The lower part of the graph is relevant only
        for periods where rapid tests are available.}
\end{figure}

In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
possible infection progresses as shown in Figure~\ref{fig:model_disease_progression}. We
differentiate between an initial period of infection without being infectious or showing
symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
intensive care, and recovery or death \citep[similar
to][]{Grimm2021}.\comment[id=HM]{Need to look for non-economists} The probabilities of
transitioning between these states depend on age; their duration is random within
intervals calibrated to medical literature (for a detailed description see
Section~\ref{sec:medical_params}). Conditional on the type of contact, infectiousness is
independent of age \citep{Jones2021}.\comment[id=K]{This might be misleading because the
susceptibility to catch the virus \textbf{is} age-dependent.}

@@

We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
individual is infected or not; there is no uncertainty to the result. PCR tests require
some days to be processed and there are aggregate capacity constraints throughout. In
contrast, rapid antigen tests yield immediate results; after a phase-in period, all tests
that are demanded will be performed.\comment[id=K]{Might be problematic to talk about
supply and demand here. We don't have supply and demand (except for maybe employers and
employees) and I would not interpret the increases in April as an easing of supply side
constraints.} Specificity and sensitivity of these tests is set according to data
analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on the timing of the test
relative to the onset of infectiousness. Figure~\ref{fig:model_rapid_tests} shows our
model for rapid test demand. Schools may require staff and students to be tested
regularly. Rapid tests may be offered by employers to on-site workers. Individuals may
demand tests for private reasons, which include having plans to meet other people,
showing symptoms of CoViD-19, and because a household member tested positively for the
virus. We endow each agent with an individual compliance parameter. This parameter
determines whether she takes up rapid tests\deleted[id=K]{offered by employers or
follows up on private reasons} % compliance also determines who is tested in education settings
.\footnote{A positive test result (of either kind), as well as symptoms leads most
individuals to reduce their contacts; this is why tests impact the actual contacts in
Figure~\ref{fig:model_description}.}

%%% ===================================================================================

@@

The main issue with official case numbers is that they will contain only a fraction of
all infections. We thus model official cases as depicted in
Figure~\ref{fig:model_official_cases}.\comment[id=K]{I don't understand why we mention
the false positive rates in the floatfoot of this. I would either mention rapid test
sensitivity and specificity in general or not mention them at all.} We take aggregate
estimates of the share of detected cases and use data on whether CoViD-19 symptoms led to
a PCR test. As the share of asymptomatic individuals varies by age group, this gives us
age-specific shares (see Figure~\ref{fig:share_known_cases_by_age_group} for the share of
known cases by age group over time in our model). Our estimates suggest that---in the
absence of rapid testing---the detection rate is on average 80\% higher for individuals
above age 80 compared to school age children. Once rapid test become available,
confirmation of a positive result is another reason leading to positive PCR tests.

@@

    \floatfoot{\noindent \textit{Note:} Data sources are described in Supplementary
        Material~\ref{sec:data_and_parameters}. Age- and region-specific analogues to
        Figure~\ref{fig:aggregated_fit} can be found in Supplementary Material
        \ref{subsec:fit_results}. For legibility reasons, all lines in
        Figure~\ref{fig:stringency_infectious_contacts} are rolling 7-day averages. The
        Oxford Response Stringency Index is scaled as $2 \cdot (1 -  x / 100)$, so that
        a value of 1 is refers to the situation at the start of our sample period and 0
        means that all NPIs included in the index are turned on. The other lines in
        Figure~\ref{fig:stringency_infectious_contacts} show the product of the effect
        of contact reductions, increased hygiene regulations, and seasonality. See
        Appendix~\ref{subsec:policies_seasonality} for separate plots of the three
        factors by contact type.}
\end{figure}

@@

Until mid-March, there is no visible difference between the different scenarios.
Seasonality hardly changes, and only few vaccinations or rapid tests were administered.
Even thereafter, the effect of the vaccination campaign is surprisingly small at first
sight. Whether considering recorded or total infections with only one channel active,
the final level is always the highest in case of the vaccination campaign (orange
lines). The Shapley value decomposition shows that vaccinations contribute 17\% to the
cumulative difference between scenarios. Reasons for this are the slow start---it took
until March~24th until 10\% of the population had received their first vaccination, the
20\% mark was reached on April 19th---and the focus on older individuals. These groups
contribute less to the spread of the disease than others due to a lower number of
contacts, see Supplementary Material~\ref{subsec:contacts_by_age}. By the end of our
study period, when first-dose vaccination rates reached around 40\% of the population,
the numbers of new cases would have started to decline. It is important to note that the
initial focus of the campaign was to prevent deaths and severe disease. Indeed, the case
fatality was rate considerably lower during the third wave when compared to the second
(4.4\% between October and February and 1.4\% between March and the end of May).

Seasonality has a large effect in slowing the spread of SARS-CoV-2. By May 31, both
observed and recorded cases would be reduced by a factor of four if only seasonality
mattered. However, in this period, cases would have kept on rising throughout, just at a
much lower pace \citep[this is in line with results in][]{Gavenciak2021}. Nevertheless,
we estimate seasonality to be a quantitatively important factor determining the
evolution of the pandemic, explaining most of the early changes and 38\% of the
cumulative difference by the end of May.

The largest effect---45\% in the decompositions---comes from rapid testing. Here, it is
crucial to differentiate between recorded cases and actual cases. Additional testing
means that additional infections will be recorded which would otherwise remain
undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is large.
Until late April, recorded cases are higher in the scenario with rapid testing alone
when compared to the setting where none of the three mechanisms are turned on. Compared
to the scenario with vaccinations only, this point is reached around mid-May and it
would be June for the comparison with the seasonality-only scenario. The effect on total
cases, however, is visible immediately in
Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only 10\%  of the
population performed weekly rapid tests in March on average, new infections on April 1
would be reduced by 53\% relative to the scenario without vaccinations, rapid tests, or
seasonality.

So why is rapid testing so effective? In order to shed more light on this question,
Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
scenario without rapid tests only (purple line in
Figure~\ref{fig:2021_scenarios_newly_infected}) and the main specification into the
three channels for rapid tests. Tests at schools have the smallest effect, which is
largely explained by schools not operating at full capacity during our period of study
and the relatively small number of students.\footnote{18\% of our population are in the
education sector (pupils, teachers etc.); 46\% are workers outside the education sector}
Almost 40\% come from tests at the workplace. Despite the fact that rapid tests for
private reasons are phased in only late, they make up for more than half of the total
effect. The reason lies in the fact that a substantial share of these tests is driven by
an elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or
following up on a positive test of a household member. The latter is essentially a form
of contact tracing, which has been shown to be very effective \citep{Contreras2021,
Kretzschmar2020,FetzerGraeber2021}.

Two of the most contentious NPIs concern schools and mandates to work from home. In many
countries, schools switched to remote instruction during the first wave, so did Germany.
After the summer break, they were operating at full capacity with increased hygiene
measures, before being closed again from mid-December onwards. Some states started
opening them gradually in late February, but operation at normal capacity did not resume
until the beginning of June. Figure~\ref{fig:school_scenarios} shows the effects of
different policies regarding school starting at Easter, at which point rapid tests had
become widely available. We estimate the realized scenario to have essentially the same
effect as a situation with closed schools. Under fully opened schools with mandatory
tests, total infections would have been 7\% higher; this number rises to 20\% without
tests. These effect sizes are broadly in line with empirical studies, e.g.,
\citep{Vlachos2021}\comment[id=HM]{Need a back-of the envelope calculation}
To use another metric, the effective weekly reproduction number differs by $0.017$ and
$0.045$, respectively. In light of the large negative effects school closures have on
children and parents \citep{Luijten2021, Melegari2021}---and in particular on those with
low socio-economic status---these results in conjunction with hindsight bias suggest that
opening schools combined with a testing strategy would have been beneficial. In other
situations, and particular when rapid test are not available at scale, trade-offs may
well be different.

@@

Figure~\ref{fig:workplace_scenarios} shows that with a large fraction of workers
receiving tests, testing at the workplace has larger effects than mandating employees to
work from home. Whether the share of workers working at the usual workplace is reduced
or increased by ten percent hardly has an effect on infection rates. Making testing
mandatory twice a week---assuming independent compliance by employers and workers or
95\% each---would have reduced infections by 13\%. Reducing rapid tests offers by
employers to the level of March would have increased infections by the same amount.

>>>>

This paper develops a quantitative model incorporating these factors simultaneously. The
framework allows to combine a wide variety of data and mechanisms in a timely fashion,
making it useful to predict the effects of various interventions. We apply the model to
Germany, where new infections fell by almost 80\% during the month of May 2021. Our
analysis shows that, aside from seasonality, frequent and large-scale rapid testing
caused the bulk of this decrease, which is in line with prior predictions
\citep{Mina2021}. We conclude that it should have a large role for at least as long as
vaccinations have not been offered to an entire population.

At the core of our agent-based model are physical contacts between heterogeneous agents
(Figure~\ref{fig:model_contacts_infections}).\footnote{A detailed comparison with other
approaches  is relegated to Supplementary Material~\ref{sec:literature_review}. The
model most closely related to ours is described in \citet{Hinch2020}.} Each contact
between an infectious individual and somebody susceptible to the disease bears the risk
of transmitting the virus. Contacts occur in up to four networks: Within the household,
at work, at school, or in other settings (leisure activities, grocery shopping, medical
appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
applications can take the population and household structure from census data and the
network-specific frequencies of contacts from diary data measuring contacts before the
pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies
depend on age and geographical location (see Supplementary Material
Section~\ref{sub:number_of_contacts}).

@@

    \floatfoot{\noindent \textit{Note:} A% formal
        description of the model can be found in Supplementary Material~\ref{sec:model}.
        Figure~\ref{fig:model_contacts_infections} shows the influence on an agent's
        contacts to other agents. Demographic characteristics set the baseline number of
        contacts in different networks. She may reduce the number of contacts due to
        NPIs, showing symptoms, or testing positively for SARS-CoV-2. Infections may
        occur when a susceptible agent meets an infectious agent; the probability
        depends on the type of contact, on seasonality, and on NPIs. If infected, the
        infection progresses as depicted in Figure~\ref{fig:model_disease_progression}.
        If rapid tests are available, agents' demand is modeled as in
        Figure~\ref{fig:model_rapid_tests}. All reasons trigger a test only for a
        fraction of individuals depending on an individual compliance parameter; the
        thresholds for triggering test demand differ across reasons and they may depend
        on calendar time. Figure~\ref{fig:model_official_cases} shows the model of
        translating all infections in the simulated data to age-specific recorded
        infections. The model uses data on the aggregate share of recorded cases, on the
        share of positive PCR tests triggered by symptoms, and on the false positive
        rate of rapid tests. The lower part of the graph is relevant only for periods
        where rapid tests are available.}
\end{figure}

In the model, susceptibility to contracting the SARS-CoV-2 virus is dependent on age. A
possible infection progresses as shown in Figure~\ref{fig:model_disease_progression}. We
differentiate between an initial period of infection without being infectious or showing
symptoms, being infectious (presymptomatic or asymptomatic), showing symptoms, requiring
intensive care, and recovery or death \citep[similar to][]{Grimm2021}.
\comment[id=HM]{Need to look for non-economists} The probabilities of transitioning
between these states depend on age; their duration is random within intervals calibrated
to medical literature (for a detailed description see Section~\ref{sec:medical_params}).
Conditional on the type of contact, infectiousness is independent of age
\citep{Jones2021}.

@@

We include two types of tests. Polymerase chain reaction (PCR) tests reveal whether an
individual is infected or not; there is no uncertainty to the result. PCR tests require
some days to be processed and there are aggregate capacity constraints throughout. In
contrast, rapid antigen tests yield immediate results; after a phase-in period, all
tests that are demanded will be performed. Specificity and sensitivity of these tests is
set according to data analyzed in \cite{Bruemmer2021, Smith2021}; sensitivity depends on
the timing of the test relative to the onset of infectiousness.
Figure~\ref{fig:model_rapid_tests} shows our model for rapid test demand. Schools may
require staff and students to be tested regularly. Rapid tests may be offered by
employers to on-site workers. Individuals may demand tests for private reasons, which
include having plans to meet other people, showing symptoms of CoViD-19, and because a
household member tested positively for the virus. We endow each agent with an individual
compliance parameter. This parameter determines whether she takes up rapid
tests.\footnote{A positive test result (of either kind), as well as symptoms leads most
individuals to reduce their contacts; this is why tests impact the actual contacts in
Figure~\ref{fig:model_description}.}

@@

The main issue with official case numbers is that they will contain only a fraction of
all infections. We thus model official cases as depicted in
Figure~\ref{fig:model_official_cases}. We take aggregate estimates of the share of
detected cases and use data on whether CoViD-19 symptoms led to a PCR test. As the share
of asymptomatic individuals varies by age group, this gives us age-specific shares (see
Figure~\ref{fig:share_known_cases_by_age_group} for the share of known cases by age
group over time in our model). Our estimates suggest that---in the absence of rapid
testing---the detection rate is 80\% higher on average for individuals above age 80
compared to school age children. Once rapid test become available, confirmation of a
positive result is another reason leading to positive PCR tests.

@@

    \floatfoot{\noindent \textit{Note:} Data sources are described in Supplementary
        Material~\ref{sec:data_and_parameters}. Age- and region-specific analogues to
        Figure~\ref{fig:aggregated_fit} can be found in Supplementary Material
        \ref{subsec:fit_results}. For legibility reasons, all lines in
        Figure~\ref{fig:stringency_infectious_contacts} are rolling 7-day averages. The
        Oxford Response Stringency Index is scaled as $2 \cdot (1 -  x / 100)$, so that
        a value of 1 refers to the situation at the start of our sample period and 0
        means that all NPIs included in the index are turned on. The other lines in
        Figure~\ref{fig:stringency_infectious_contacts} show the product of the effect
        of contact reductions, increased hygiene regulations, and seasonality. See
        Appendix~\ref{subsec:policies_seasonality} for separate plots of the three
        factors by contact type.}
\end{figure}

@@

Until mid-March, there is no visible difference between the different scenarios.
Seasonality hardly changes, and only few vaccinations or rapid tests were administered.
Even thereafter, the effect of the vaccination campaign is surprisingly small at first
sight. Whether considering recorded or total infections with only one channel active,
the final level is always the highest in case of the vaccination campaign (orange
lines). The Shapley value decomposition shows that vaccinations contribute 16\% to the
cumulative difference between scenarios. Reasons for this are the slow start---it took
until March~24th until 10\% of the population had received their first vaccination, the
20\% mark was reached on April 19th---and the focus on older individuals. These groups
contribute less to the spread of the disease than others due to a lower number of
contacts, see Supplementary Material~\ref{subsec:contacts_by_age}. By the end of our
study period, when first-dose vaccination rates reached around 40\% of the population,
the numbers of new cases would have started to decline. It is important to note that the
initial focus of the campaign was to prevent deaths and severe disease. Indeed, the case
fatality was rate considerably lower during the third wave when compared to the second
(4.4\% between October and February and 1.4\% between March and the end of May).

Seasonality has a large effect in slowing the spread of SARS-CoV-2. By May 31, both
observed and recorded cases would be reduced by a factor of four if only seasonality
mattered. However, in this period, cases would have kept on rising throughout, just at a
much lower pace \citep[this is in line with results in][, which our seasonality measure
is based on]{Gavenciak2021}. Nevertheless, we estimate seasonality to be a
quantitatively important factor determining the evolution of the pandemic, explaining
most of the early changes and 43\% of the cumulative difference by the end of May.

A similar-sized effect---41\% in the decompositions---comes from rapid testing. Here, it
is crucial to differentiate between recorded cases and actual cases. Additional testing
means that additional infections will be recorded which would otherwise remain
undetected. Figure~\ref{fig:2021_scenarios_recorded} shows that this effect is large.
Until late April, recorded cases are higher in the scenario with rapid testing alone
when compared to the setting where none of the three mechanisms are turned on. The
effect on total cases, however, is visible immediately in
Figure~\ref{fig:2021_scenarios_newly_infected}. Despite the fact that only 10\%  of the
population performed weekly rapid tests in March on average, new infections on April 1
would be reduced by 53\% relative to the scenario without vaccinations, rapid tests, or
seasonality.

So why is rapid testing so effective? In order to shed more light on this question,
Figure~\ref{fig:2021_scenarios_decomposition_tests} decomposes the difference in the
scenario without rapid tests only (purple line in
Figure~\ref{fig:2021_scenarios_newly_infected}) and the main specification into the
three channels for rapid tests. Tests at schools have the smallest effect, which is
largely explained by schools not operating at full capacity during our period of study
and the relatively small number of students.\footnote{18\% of our population are in the
education sector (pupils, teachers, etc.); 46\% are workers outside the education sector}
Almost 40\% come from tests at the workplace. Despite the fact that rapid tests for
private reasons are phased in only late, they make up for more than half of the total
effect. The reason lies in the fact that a substantial share of these tests is driven by
an elevated probability to carry the virus, i.e., showing symptoms of CoViD-19 or
following up on a positive test of a household member. The latter is essentially a form
of contact tracing, which has been shown to be very effective \citep{Contreras2021,
Kretzschmar2020,FetzerGraeber2021}.

Two of the most contentious NPIs concern schools and mandates to work from home. In many
countries, schools switched to remote instruction during the first wave, so did Germany.
After the summer break, they were operating at full capacity with increased hygiene
measures, before being closed again from mid-December onwards. Some states started
opening them gradually in late February, but operation at normal capacity did not resume
until the beginning of June. Figure~\ref{fig:school_scenarios} shows the effects of
different policies regarding school starting at Easter, at which point rapid tests had
become widely available. We estimate the realized scenario to have essentially the same
effect as a situation with closed schools. Under fully opened schools with mandatory
tests, total infections would have been 6\% higher; this number rises to 20\% without
tests. These effect sizes are broadly in line with empirical studies, e.g.,
\citep{Vlachos2021}\comment[id=HM]{Need a back-of the envelope calculation} To use
another metric, the effective weekly reproduction number differs by $0.018$ and $0.052$,
respectively. In light of the large negative effects school closures have on children
and parents \citep{Luijten2021, Melegari2021}---and in particular on those with low
socio-economic status---these results in conjunction with hindsight bias suggest that
opening schools combined with a testing strategy would have been beneficial. In other
situations, and particular when rapid test are not available at scale, trade-offs may
well be different.

@@

Figure~\ref{fig:workplace_scenarios} shows that with a large fraction of workers
receiving tests, testing at the workplace has larger effects than mandating employees to
work from home. Whether the share of workers working at the usual workplace is reduced
or increased by ten percent changes infection rates by 2.5\% or less in either
direction. Making testing mandatory twice a week---assuming independent compliance by
employers and workers of 95\% each---would have reduced infections by 23\%. Reducing
rapid tests offers by employers to the level of March would have increased infections by
13\%.
